Optimizing the power of adaptive seamless designs using Bayesian Optimization

### 2020-12-09 main.tex
besprechung 9.12.2020

The increasing complexity of both the designs and the analyses of clinical trials has the consequence that the planning of such studies often relies also on computer simulations, especially if no closed form solutions for calculations are available.
This applies even to usually fairly straightforward tasks as sample size calculations. 
Various frameworks and guidance proposals have been developed over the past years on how to plan, execute and interpret the results of such simulation studies.
Method comparison studies are used to give clinicians and biostatisticians guidance when an approach improves currently used methods.  
Depending on the setting and the purpose of the simulation study, suitable metrics must be chosen to compare alternative designs or analysis strategies. 
Especially relevant are neutral comparison studies, which are characterized as follows:
%FIXME Is there someting missing? (Jakob)
The primary goal of the respective article is not to introduce a new promising method, the authors are reasonably neutral, and the evaluation metrics, methods, and data sets are chosen in a rational way~\citep{boulesteix_plea_2013}.
%While method comparison studies aim to make statements when one approach might dominate others in terms of the chosen metrics, in designing an individual study or series of studies the objective is to identify a design that is optimal in terms of the metrics chosen in a certain scenario or, more importantly, across a set of scenarios. 
A comprehensive framework for such an objective assessment of competing strategies for clinical trial designs is CSE (Clinical Scenario Evaluation). ref %FIXME ref
CSE was particularly developed to support the overall process of performing simulations of clinical trials, especially for complex designs and analysis strategies.
The problem is decomposed into data models, analysis models, and evaluation models that specify the evaluation metrics.

@@

The proposal of a new configuration $\theta$ is obtained by maximizing a so called acquisition function.
The acquisition function guides the search to promising new regions in the search space $\Theta$ by facilitating the mean prediction $\hat{\mu}(\theta)$ as well as the uncertainty quantification $s^2(\theta)$ of the surrogate.
It balances between exploration of not yet evaluated regions in $\Theta$ and exploitation.
The acquisition function achieves exploration by assigning high values to areas where the surrogate predicts a high uncertainty.
Exploitation is achieved by assigning high values to areas of $\theta$ where the surrogate predicts a high outcome of $f$.
For deterministic functions the expected improvement~\citep{jones_efficient_1998} is arguably the most popular acquisition function.

>>>>

The increasing complexity of both the designs and the analyses of clinical trials has the consequence that the planning of such studies often relies also on computer simulations, especially if no closed form solutions for calculations are available.
This applies even to usually fairly straightforward tasks as sample size calculations. 
Various frameworks and guidance proposals have been developed over the past years on how to plan, execute and interpret the results of such simulation studies.
Method comparison studies are used to give clinicians and biostatisticians guidance when an approach improves currently used methods.  
Depending on the setting and the purpose of the simulation study, suitable metrics must be chosen to compare alternative designs or analysis strategies. 
Especially relevant are neutral comparison studies, which are characterized by~\citet{boulesteix_plea_2013} as follows:
%FIXME Is there someting missing? (Jakob)
The primary goal of the respective article is not to introduce a new promising method, the authors are reasonably neutral, and the evaluation metrics, methods, and data sets are chosen in a rational way.
%While method comparison studies aim to make statements when one approach might dominate others in terms of the chosen metrics, in designing an individual study or series of studies the objective is to identify a design that is optimal in terms of the metrics chosen in a certain scenario or, more importantly, across a set of scenarios. 
A comprehensive framework for such an objective assessment of competing strategies for clinical trial designs is CSE (Clinical Scenario Evaluation)~\citep{friede_refinement_2010}.
CSE was particularly developed to support the overall process of performing simulations of clinical trials, especially for complex designs and analysis strategies.
The problem is decomposed into data models, analysis models, and evaluation models that specify the evaluation metrics.

@@

The proposal of a new configuration $\theta$ is obtained by maximizing a so called acquisition function.
The acquisition function guides the search to promising new regions in the search space $\Theta$ by using the mean prediction $\hat{\mu}(\theta)$ as well as the uncertainty quantification $s^2(\theta)$ of the surrogate.
It balances between exploration of not yet evaluated regions in $\Theta$ and exploitation.
The acquisition function achieves exploration by assigning high values to areas where the surrogate predicts a high uncertainty.
Exploitation is achieved by assigning high values to areas of $\theta$ where the surrogate predicts a high outcome of $f$.
For deterministic functions the expected improvement~\citep{jones_efficient_1998} is arguably the most popular acquisition function.
